---
title: "Going the distance_test"
subtitle: "The effects of commute distance on teacher recruitment difficulties"
author: "Rasmus H. Klokker"
institute: "VIVE"
date: "29/04/2024 (updated: `r Sys.Date()`)"
editor_options:
  chunk_output_type: console
output:
  xaringan::moon_reader:
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg)


```{r setup, echo = FALSE, warning = FALSE, message=FALSE}



options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, dev = "png", dev.args = list(type = "cairo-png"))

#knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, dev = "svglite")




library(ggmap)
library(RCurl)

library(httr)
library(jsonlite)
library(plyr)
library(patchwork)
library(gridExtra)
library(geosphere)
library(scales)
library(kableExtra)
library(purrr)
library(rgdal)
library(plotly)

#setwd("C:\\Users\\B059064\\Desktop\\PHD\\conferences and presentations\\CEN 2023")

sogn_travel_dist <- read.table("tables\\sogn_travel_dist.txt",sep=",",check.names=FALSE)


ps <- read.table("tables\\des_stats_covars.txt",sep=",",check.names=FALSE)

ps_tchrs <- read.table("tables\\des_stats_tchrs.txt", check.names=FALSE, sep=",")


ps_studs <- read.table("tables\\des_stats_studs.txt",sep=",",check.names=FALSE)


res_df <- read.table("tables\\DID_commute.txt",sep=",",check.names=FALSE)


did_estimaes_df_tchrs_red <- read.table("tables\\did_estimaes_tchrs.txt",sep=",",check.names=FALSE)


did_estimaes_df_studs_red <- read.table("tables\\did_estimaes_studs.txt",sep=",",check.names=FALSE)


effs <- read.table("tables\\effect_sizes_thresholds.txt",sep=",",check.names=FALSE)



effs_studs <- read.table("tables\\effect_sizes_thresholds_studs.txt",sep=",",check.names=FALSE)


# alt_estimators_tchrs <- readRDS("alt_estimators_tchrs.rds")

did_estimaes_df_tchrs <- read.table("tables\\did_estimaes_tchrs_full.txt",sep=",",check.names=FALSE)



# alt_estimators_studs <- readRDS("alt_estimators_studs.rds")

did_estimaes_df_studs <- read.table("tables\\did_estimaes_studs_full.txt",sep=",",check.names=FALSE)

dist_thresh_dfs <- read.table("C:\\Users\\B059064\\Desktop\\PHD\\papers\\TTC dif in dif\\PHD_dissertation\\from_DST_results\\results_5-3-2024\\tables\\att_across_dist_thresholds_tchrs.txt", sep=",", check.names=FALSE)


bsadid_df_all_comb_mean_tchrs <- read.table("C:\\Users\\B059064\\Desktop\\PHD\\papers\\TTC dif in dif\\PHD_dissertation\\from_DST_results\\results_5-3-2024\\tables\\att_cohort_tchrs.txt", sep=",",check.names=FALSE)



access_df_agg <- read.table("tables\\tchr_access_df.txt",sep=",",check.names=FALSE)





access_df_agg_2 <- read.table("tables\\tchr_access_df_cohort.txt",sep=",",check.names=FALSE)



access_df_agg_2$dif <- access_df_agg_2$tcrh_access - access_df_agg_2$tchr_access_year


# 
# 
# 
# df <- read.table("tables\\number_grads_by_ttc.txt",sep=",",check.names=FALSE)
# 
# 
# df_2 <- read.table("tables\\number_teachers_by_experXschool_reduced.txt",sep=",",check.names=FALSE)
# 


# bsadid_df_all_comb_mean_tchrs <- read.table("tables\\att_cohort_tchrs.txt", sep=",",check.names=FALSE)
# 
# bsadid_df_all_comb_mean_studs <- read.table("tables\\att_cohort_studs.txt", sep=",",check.names=FALSE)


alt_estimators_df_tchrs <- read.table("tables\\alt_estimators_tchrs.txt", sep=",", 
                                      check.names = FALSE)

alt_estimators_cohort_df_tchrs <- read.table("tables\\alt_estimators_cohorts_tchrs.txt", sep=",", check.names = FALSE)

alt_estimators_df_studs <- read.table("tables\\alt_estimators_studs.txt", sep=",", 
                                      check.names = FALSE)

alt_estimators_cohort_df_studs <- read.table("tables\\alt_estimators_cohort_studs.txt", sep=",", check.names = FALSE)


dist_thresh_dfs <- read.table("tables\\att_across_dist_thresholds_tchrs.txt", sep=",", check.names=FALSE)

dist_thresh_dfs_stud <- read.table("tables\\att_across_dist_thresholds_studs.txt", sep=",", check.names=FALSE)

outcomes_list_tchrs <- c("avg_age", "avg_seniority", "tchr_pupil_ratio", "teacher_ed_share_new", "avg_hs_GPA", "avg_hs_GPA_young", "avg_year_since_grad", "newgrad_teacher_share")

outcomes_list_studs <- c("avg_hs_GPA_students", "pria_5", "pria_10", "pria_20", "secondedcomp_5", "secondedcomp_10", "secondedcomp_20", "collegecomp_10", "collegecomp_15", "collegecomp_20")


#ggplot theme

theme_set(theme_minimal(base_size = 16,base_family = "serif")+ theme(axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()))


```


```{r xaringan-fit-screen, echo=FALSE}
xaringanExtra::use_fit_screen()
```




```{r set up TTC data, include=FALSE}


#load TTC data------

#only run these lines if you haven't gotten the geodata for the TTC's yet

# ttc.info <- read.csv("TTC_info.csv",sep=";", encoding = "latin1")
# 
# #Query Google Maps API for TTC coordinates----
# 
# 
# 
# 
# 
# ##get geodata loop-------
# 
# geodata.dfs <- lapply(1:nrow(ttc.info), function(x) {
#   
# cat(paste("processing row nr.", x, "\r"))  #progress indicator
#   
#   ID <- x
# 
# #format address for the API query  
#   
# address <- trimws(gsub(" ", "%20", ttc.info$adresse[x])) #insert %20 instead of white space
# 
# address <- gsub("\n", "", address) #get rid of line breaks
# 
# #construct API query
# 
# query <- paste("https://nominatim.openstreetmap.org/search/", address, "?format=json&polygon=1&addressdetails=1", sep="")
# 
# #query API
# 
# coords <- httr::GET(query)
# 
# #we get a json response, but it's in binary code, so we'll reformat using the "content" function from "httr"
# 
# json<-content(coords,as="text") 
# 
# #and finally we can get our dataframe for the wuery in question
# 
# geodata <- fromJSON(json)
# 
# #The dataframe has a couple of problematic columns that we'll need to deal with, since we need
# #to rbind all the dataframes which we get from the loop
# 
# 
# #firstly the "address" column is a dataframe, so well make that into a separate datafram and cbind it to out main dataframe
# addr.df <- geodata$address
# 
# geodata <- cbind(geodata[,-ncol(geodata)], addr.df)
# 
# #the column "boundingbox" is a list. Since we don't really need this column, we'll just remove it from the dataframe
# 
# ind <- which(names(geodata) %in% "boundingbox")
# 
# geodata <- geodata[,-ind]
# 
# geodata$ttc_id <- ID
# 
# return(geodata)
# 
# })
# 
# 
# 
# geodata.full <- do.call("rbind.fill", geodata.dfs)
# 
# #remove duplicates
# 
# geodata.full <- geodata.full[!duplicated(geodata.full[,"ttc_id"]),]
# 
# 
# geodata.full <- cbind(geodata.full, ttc.info)
# 
# geodata.full$år.nedlagt[is.na(geodata.full$år.nedlagt)] <- 9999
# 
#write.csv(geodata.full, file="geodata_TTC.csv")

geodata.full <- read.csv("geodata_TTC.csv", fileEncoding="latin1")

#names(geodata.full)[grep("Ã.r.nedlagt$", names(geodata.full))] <- "year_closed"





schools.info <- read.csv("C:\\Users\\B059064\\Desktop\\PHD\\papers\\TTC dif in dif\\PHD_dissertation\\folkeskoledata_all_corrected.csv")


#compute distance between schools and TTC's--------


schools.info$GEO_BREDDE_GRAD <- as.numeric(gsub(",", ".", schools.info$GEO_BREDDE_GRAD))

schools.info$GEO_LAENGDE_GRAD <- as.numeric(gsub(",", ".", schools.info$GEO_LAENGDE_GRAD))

shp <- readOGR(dsn = "C:\\Users\\B059064\\Desktop\\PHD\\papers\\TTC dif in dif\\PHD_dissertation\\DK_shp\\DNK_adm0.shp", 
               stringsAsFactors = F)


```





# Introduction

- Teachers are not distributed equally across schools! 

- Worse, the best teachers are often employed at the schools with the most well off students(Nguyen, 2020; Guarino et al., 2006; Borman & Dowling, 2008)

- Unequal sorting of teachers excacerbate achievement gap between low- and high-SES students(Boyd et al., 2005; Goldhaber, 2012; Sass et al., 2012)

--

- *But how do different characteristics of schools impact how well they can attract teachers?*


- *Can we alter characteristics of schools to promote a more equal distribution of teachers?*


---

# This paper

- In this paper i investigate what happens when *commute distance* increases for teachers, _specifically newly graduated teachers_.

- To do this, i investigate a natural experiment in which over 33% of danish teacher colleges(TC's) closed

- This meant that some schools experienced an increase in the distance to the nearest TC

## Why commute distance?

- Potential policy lever


---


# Previous research

## Teacher sorting is a thing!

- The literature generally shows that teachers are unequally distributed across schools

- Both teacher and schools traits impact how teachers are distributed across schools

- teacher credentials(Boyd et al., 2005a; Lankford et al.,
2007; Podgursky et al., 2004)

- school level poverty(Allen et al., 2018; Smith and Ingersoll, 2007), school
level achievement(Hanushek et al., 2015; Ingersoll and May, 2012)

- school level achievement(Hanushek et al., 2015; Ingersoll and May, 2012), higher proportions of ethnic minority
pupils(Allen et al., 2018; Carroll et al., 2000; Falch and Strøm, 2004; Feng, 2014; Hanushek et al., 2015;
Shen, 1997)

- higher proportions of ethnic minority
pupils(Allen et al., 2018; Carroll et al., 2000; Falch and Strøm, 2004; Feng, 2014; Hanushek et al., 2015;
Shen, 1997)

- higher shares of inexperienced teachers, discipline problems, and inadequate support from the
administration(Ingersoll, 2007a; Shen, 1997)

---

# Previous research


## Is teacher sorting such a big deal

- teacher turnover harms student
achievement(Gibbons, 2018; Henry and Redding, 2020)

- highly skilled
teachers are more responsive to changes in school characteristics that might change the attractiveness of
schools, e.g., student body composition(Goldhaber, 2010; Jackson, 2009)

- Studies also find that disadvantaged
schools “inherit” low-quality teachers from other disadvantaged schools, while high-quality teachers
transfer to advantaged schools(Goldhaber 2012; Sass et al., 2012)

- Newly graduated highly skilled teachers select into advantaged
schools(Krieg et al., 2015)


---

# Previous research

## Does commute distance seem to impact school attractiveness?

- Commute distance’s impact on teacher attraction has been overlooked compared to other
factors like student body composition. 

- increasing commute distance from 0 to 6 miles leads to a 19% increase in the probability of
transferring to a different school(Boyd et al., 2005a)

- living within a distance of 1 mile from an average urban school still preferable to working at an average suburban school located at least 21 miles away. This is despite there being a 60 and 52 percentage point difference between
urban and suburban schools in proportion of minority students and students living in poverty, respectively,
as well as urban schools offering lower salaries (Boyd et al., 2012) 

- 10 minutes commute equivalent to a reduction in yearly salary of 530$. By comparison, a ten-percentage point
reduction in the proportion of students living in poverty is equivalent to a 320$ increase in salary and a
ten-point increase in the average percentile of performance is equivalent to an increase in salary of 550$ (Johnston, 2020)

---

## How about distance to TTC?


- New teachers are 36% more likely to find their first teaching position in the region in which they
obtained their last educational credential, compared to other regions (Boyd et al., 2005b)

- Prospect teachers complete
their student teaching near the college from which they obtained their degree, and that student teaching
placement is highly predictive of where newly graduated teachers find their first teaching position (Krieg et al., 2015)

- An increase in distance to the nearest TTC results in a
higher proportion of newly hired teachers (Goldhaber et al., 2020)

- number of applicants for vacant positions decreases as the distance to the nearest
TTC increases (Goff and Bruecker, 2017)

- The relation between the proximity of schools to TTCs and teacher recruitment has only
been addressed in a few studies(Boyd et al.,2005a)

- however results show that distance to the nearest TTC likey
has an impact on teacher recruitment(Boyd et al., 2005a; Goldhaber et al., 2020; Krieg et
al., 2015) Goff and Bruecker (2017)




---



# This paper

## Hypothesis

```{r hypo-anim}



# schools.info <- read.csv("C:\\Users\\B059064\\Desktop\\PHD\\papers\\TTC dif in dif\\PHD_dissertation\\folkeskoledata_all_corrected.csv")


# geodata.full <- read.csv("C:\\Users\\B059064\\Desktop\\PHD\\papers\\TTC dif in dif\\PHD_dissertation\\geodata_TTC.csv")
# 
# geodata_full_samp <- geodata.full[sample(1:nrow(geodata.full), 5),]

pattern <- "Frederiksberg|Haderslev Seminarium|Esbjerg|Odense|Silkeborg"



geodata_full_samp <- geodata.full[grep(pattern, geodata.full$Seminarium),]


#names(geodata.full)[grep("Ã.r.nedlagt$", names(geodata.full))] <- "year_closed"


ttc_radius_df <- purrr::map_dfr(1:nrow(geodata_full_samp), function(x){
  
  
  point <- sf::st_point(c(geodata_full_samp$lat[x], geodata_full_samp$lon[x]))
  
  point <- sf::st_sfc(point, crs="+proj=longlat +datum=WGS84 +no_defs")
  
  buf_round <- sf::st_buffer(point, dist = 20000)
  
  buf_round <-as.data.frame(sf::st_coordinates(buf_round))
  
  buf_round$ttc <- x
  
  return(buf_round)
  

  
  
})



#compute distance between schools and TTC's--------


# schools.info$GEO_BREDDE_GRAD <- as.numeric(gsub(",", ".", schools.info$GEO_BREDDE_GRAD))
# 
# schools.info$GEO_LAENGDE_GRAD <- as.numeric(gsub(",", ".", schools.info$GEO_LAENGDE_GRAD))

set.seed(43920)

schools.info_samp <- schools.info[sample(1:nrow(schools.info), 20),]

tch_df <- data.frame(tch_id=1:200, lat=0, lon=0)

# tch_df$lon <- rnorm(200,mean(c(shp@bbox[1],shp@bbox[3])))
#                     
# 
# tch_df$lat <- rnorm(200,mean(c(shp@bbox[2],shp@bbox[4])))


#simply take coordinates from school coordinates, to avoid teachers randomluy landing in the sea


set.seed(43920)


tch_df[c("lat", "lon")] <- schools.info[sample(1:nrow(schools.info),200, replace=TRUE),
                                             c("GEO_BREDDE_GRAD", "GEO_LAENGDE_GRAD")]


lat_lon_pert <- purrr::map_dfr(1:max(ttc_radius_df$ttc), function(x){
  
  
  lat_lon_pert <-  schools.info[which((schools.info$GEO_BREDDE_GRAD<=max(ttc_radius_df$X[ttc_radius_df$ttc==x])&
                        schools.info$GEO_BREDDE_GRAD>=min(ttc_radius_df$X[ttc_radius_df$ttc==x]))&
                          (schools.info$GEO_LAENGDE_GRAD<=max(ttc_radius_df$Y[ttc_radius_df$ttc==x])&
                       schools.info$GEO_LAENGDE_GRAD>=min(ttc_radius_df$Y[ttc_radius_df$ttc==x]))
                       ),
               c("GEO_BREDDE_GRAD", "GEO_LAENGDE_GRAD")]
  
})

set.seed(43920)

tch_df$time <- "before studying"

tch_df_2 <- tch_df

tch_df_2[c("lat", "lon")] <- lat_lon_pert[sample(1:nrow(lat_lon_pert), 200, replace=TRUE),]

tch_df_2$time <- "while studying"

tch_df <- rbind(tch_df, tch_df_2)




dist_mat <- distm(schools.info_samp[,c("GEO_LAENGDE_GRAD", "GEO_BREDDE_GRAD")], 
                  tch_df[tch_df$time=="while studying", c("lon", "lat")], fun=distVincentyEllipsoid)

#which schools is closest to each teacher?

school_ind <- apply(dist_mat,2,which.min)


tch_df_3 <- cbind(tch_df_2$tch_id, schools.info_samp[school_ind,c("GEO_BREDDE_GRAD", "GEO_LAENGDE_GRAD")], tch_df_2$time)

names(tch_df_3) <- names(tch_df)

tch_df_3$time <- "after graduation"

tch_df <- rbind(tch_df, tch_df_3)


#make plot

p <- ggplot() + geom_polygon(data = shp, aes(x = long, y = lat, group = group), colour = "black", fill = NA)


p <- p+geom_text(data=geodata_full_samp, aes(x=lon, y=lat), inherit.aes = FALSE, size=10, label="\U1F3DB",
                 family = "Arial Unicode MS")



p_base <- p+geom_text(data=schools.info_samp, aes(x=GEO_LAENGDE_GRAD, y=GEO_BREDDE_GRAD), 
                 inherit.aes = FALSE, size=5, label="\U1F3EB", family = "Arial Unicode MS")

p_base

tch_df$time <- paste(rep(1:3, each=200), tch_df$time)


tch_df$size <- rep(c(4,6,8), each=200)

p <- p_base+geom_text(data=tch_df, aes(x=lon, y=lat, frame=time), size=3, label="\UC6C3", 
                        family = "Arial Unicode MS", position=position_jitter(width=.2,height=.2))

anim <- ggplotly(p)%>% 
  animation_opts(
    frame = 1000,
    transition = 1000,
    easing = "linear",
    redraw = FALSE,
    mode = "immediate"
  )


htmltools::save_html(anim, file="hypo_anim.html") 

#cat(plotly:::plotly_iframe(anim))

anim



```


---

<iframe src="hypo_anim.html" width="800" height="500" scrolling="yes" seamless="seamless" frameBorder="0"> </iframe>
